\begin{rubric}{Major CFD Codes Developed}
\entry*[2020 -- present]%
	\textbf{COMET:} Thermo-chemical non-Equilibrium (TCNEQ) coupled MHD CFD solver for Re-entry applications. \par
	{$\bullet$} Developed in COOLFluiD CFD platform in C++ and Python during postdoc. \par
    {$\bullet$} Coupled the TCNEQ Navier stokes equations with Maxwell equations for low magnetic Reynolds number. %\par
    %{$\bullet$} Investigate the role of magnetic and electric fields in blackout mitigation during re-entry for various atmospheres. \par
    %{$\bullet$} Investigating shock patterns and interaction of applied magnetic field with electrons in plasma with the post-shock region. \par
    %{$\bullet$} Actively using the solver for collaborating with the ground experiments at VKI, Belgium, and IRS, Germany to design of next-generation heat shields in re-usable space crafts. %\par
\entry*[2014 -- 2020]%
	\textbf{Jatayu}: In-house Reynolds Averaged Navier-Stokes (RANS) based turbulent, three-dimensional, unstructured grid, Computational Fluid Dynamics (CFD) solver for high-speed aerospace applications with the density-based algorithm. \par
    %{$\bullet$} Developed \emph{from scratch} as an objective of Ph.D. thesis. \par
    {$\bullet$} Written in C++ and parallelized using MPI with Metis for domain decomposition. \par 
    {$\bullet$} Successfully deployed on Amazon EC2 cloud computing service, with good speedups after several validation and verification studies, including full-body aircraft simulations. %\par
    %{$\bullet$} The solver is then used to conduct various fundamental cold-flow studies that can serve at least as a first approximation for the design of SCRAMJET combustors, as the second objective of the thesis.
\entry*[2013 -- 2014]%
	\textbf{Thunderstorm}: GPU parallelized 3D CFD Solver for Large Eddy Simulations. \par
	{$\bullet$} 3-D incompressible CFD unstructured grid solver for Large Eddy Simulation (LES), using Graphics Processor Units (GPU) for parallelization was developed using Nvidia CUDA and C. \par
    {$\bullet$} Coloring algorithm for domain decomposition. %\par
    %{$\bullet$} The solver was used to study the mixing and flow structures of turbulent low-speed coaxial jets using LES.
\end{rubric}